---
title: 'Peer-graded Assignment: Statistical Inference Course Project Part II'
subtitle: "Basic Inferential Data Analysis"
author: "Koen Vermeulen"
date: "10-8-2019"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

## Overview

This report will analyze tooth growth.  

## Analysis

Loading and exploring data:

```{r, echo = TRUE, results="asis"}
data("ToothGrowth")
dim(ToothGrowth)
str(ToothGrowth)
```

Basic summary:

```{r, echo = TRUE}
summary(ToothGrowth)
```

Subsetting vitamin C & orange juice:

```{r, echo=TRUE}
VC <- dplyr::filter(ToothGrowth, supp == "VC")
OJ <- dplyr::filter(ToothGrowth, supp == "OJ")
t.test(VC$len, OJ$len, paired = FALSE, var.equal = FALSE)$conf
t.test(VC$len, OJ$len, paired = FALSE, var.equal = FALSE)$p.value
```

The interval contains zero, but the p-value is above 5% so it is not conclusive.

The next step is comparing the different doses. The hypothesis is that the dose has an effect on the tooth length.

```{r, results="asis"}
dose05 <- ToothGrowth$len[ToothGrowth$dose == 0.5]
dose1 <-  ToothGrowth$len[ToothGrowth$dose == 1.0]
dose2 <-  ToothGrowth$len[ToothGrowth$dose == 2.0]
t.test(dose05, dose1)$conf
t.test(dose05, dose1)$p.value
t.test(dose05, dose2)$conf
t.test(dose05, dose2)$p.value
t.test(dose1, dose2)$conf
t.test(dose1, dose2)$p.value
```

As we can see from all 3 t.tests, the p-values are less than 0.05. Therefore we can accept the hypothesis that the dose has an effect on tooth length, in other words: the mean is not equal to zero. We also see that all of the 95% confidence intervals are below zero. This confirms that increasing the dose increases the tooth length.

In the third step we'll inspect the difference in mean lengths between the vitamin C (VC) group and orange juice (OJ) group for each dose level applied.

```{r}
dose05VC <- VC$len[VC$dose == 0.5]
dose1VC <- VC$len[VC$dose == 1.0]
dose2VC <- VC$len[VC$dose == 2.0]
dose05OJ <- OJ$len[OJ$dose == 0.5]
dose1OJ <- OJ$len[OJ$dose == 1.0]
dose2OJ <- OJ$len[OJ$dose == 2.0]
```

As in previous cases, we'll assume that observations are not paired and that variances are not equal.

```{r, results="asis"}
t.test(dose05VC, dose05OJ)$p.value
t.test(dose05VC, dose05OJ)$conf
t.test(dose1VC, dose1OJ)$p.value
t.test(dose1VC, dose1OJ)$conf
t.test(dose2VC, dose2OJ)$p.value
t.test(dose2VC, dose2OJ)$conf
```

We can see from the t-tests that the difference between group means for VC and OJ is statistically  significant for dose levels 0.5 and 1 mg/day, because the p-values are less than 0.05. The dose of 2 mg/day has a large p-value and the confidence interval includes zero, therefore there is no significant difference.

## Conclusions
The assumptions:

1. Samples used are random iid samples. 

2. Each sample is indeendent of one another, in other words, they are not paired.

3. The population distribution of each samle must be approximately normal or mound shaped and roughly symetric.

Conclusions:


**4. State your conclusions and the assumptions needed for your conclusions.**
**Assumptions:**
1. Samples used are random iid samples. 
2. Each sample is indeendent of one another, in other words, they are not paired.
3. The population distribution of each samle must be approximately normal or mound shaped and roughly symetric.
**Conclusions:**
1. Supplement type alone does not affect the mean value of length of odontoblasts (cells responsible for
toothgrowth).
2. Dose level alone, without consideraton of supplement type affects tooth growth significantly. Increasing the dosage will induce better tooth growth.
3. Orange Juice as a supplement, when used in dose levels of 0.5 mg/day and 1mg/day promotes better tooth
growth than ascorbic acid. When applied in dose level of 2 mg/day, it has an effect on tooth growth similar 
to that of an ascorbic acid.

The result of comparing all the doses is that their p-values are very low and the confidence intervals do not contain zero, so we can deny the hypothesis and conclude that the dose does affect the tooth length.

##Conclusion
In conclusion I have analysed the tooth growth data and confirmed that an increase if the dose of the supplement increases the tooth growth. It is inconclusive whether the type of supplement, vitamin C or orange juice, affects the tooth  The assumptions needed for these coclusions is that the guinea pigs were randonmly selected from a population of guinea pigs. v